{\bf {\Large Evaluation and correction of noise related to physical processes in estimates of estuary metabolism}}

### 2014-08-15 assess.Rnw
almost ready to start discussion

Multiple time series were simulated by varying the conditions in \cref{do_obs,do_bio,do_unc,do_obs_all,do_sin,do_unc_n,deltdo,deltx,do_advp,do_adv}. Specifically, the simulated data varied in the relative amount of noise in the measurement, relative amplitude of the diel \ac{DO} component, degree of association of the tide with the \ac{DO} signal, and tidal type as diurnal, semidiurnal, and mixed semidiurnal.  Three levels were evaluated for each variable: relative noise as 0, 1, and 2 standard deviations for both process and observation uncertainty, amplitude of diel biological \ac{DO} as 0, 1, and 2 mg L$^{-1}$, and \ac{DO} change from tidal advection as 0, 1, and 2 mg L$^{-1}$.  Three tidal categories were created from \cref{do_sin} using a period of 24.82 hours (principal lunar) for diurnal, 12.42 hours (principal lunar semidiurnal) for semidiurnal, and adding both diurnal and semidiurnal series for mixed semidiurnal. A total of 243 time series were created based on 81 unique combinations of parameters for each tidal category (\cref{fig:sim_ex}).  Three window widths for each variable used to estimate the total weight vetor were evaluated (as half window widths): decimal time as 1, 3, and 8 days, time of day as 6, 12, and 24 hours, and tidal height as 0.25, 0.5, and 1 as a proportion of the observed value at the center of the window.  In total, 27 window width combinations were evaluated for each of 243 simulated time series, producing results for 6561 weighted regressions.

The detided or normalized values for each regression were compared to the simulated data to evaluate the ability of weighted regression to reproduce the biological \ac{DO} signals. Results were summarized using Pearson correlation coefficients and the \ac{RMSE} between the predicted and observed \ac{DO} values and the detided and biological \ac{DO} values.  Overall, the weighted regressions provided accurate results for the detided time series compared to the `true' biological time series regardless of the simulation parameters (\cref{tab:dtd_perf1}) or window widths (\cref{tab:dtd_perf2}).  Mean correlation for all time series and window widths between the detided and biological values was \Sexpr{form_fun(perf_means['dtd_cor'])}, with values ranging from \Sexpr{form_fun(perf_rngs[1,'cor'])} to \Sexpr{form_fun(perf_rngs[2, 'cor'])}.  Mean error was \Sexpr{form_fun(perf_means['err'])}, with values ranging from \Sexpr{form_fun(perf_rngs[1,'err'])} to \Sexpr{form_fun(perf_rngs[2, 'err'])}.  Simulations with very poor performance (e.g., negative correlations) were those that had minimum widths for day windows and maximum widths for hour windows, or were those with the \ac{DO} signal composed entirely of noise from observation uncertainty. Characteristics of \ac{DO} time series that contributed to improved model performance were increasing amplitude of the diel \ac{DO} component ($DO_{die}$) and increasing process uncertainty ($e_{pro}$), whereas increasing observation uncertainty contributed to decreasing performance (\cref{tab:dtd_perf1}).  Model performance was not related to magnitude of the tidal advection component ($DO_{adv}$), whereas models with diurnal tidal components had slightly decreased performance.  Increasing widths for day and tidal proportion windows contributed to increasing model performance, whereas the opposite was true for increasing hour windows (\cref{tab:dtd_perf2}).

%%%%%
A closer examination of the results for unique scenarios can provide additional insight into characteristics of observed \ac{DO} that influence model results.  In particular, the performance of the model depends entirely on the ability to predict variation in \ac{DO} from tidal effects ($DO_{tid}$, \cref{fig:do_dtd}) and the mean response of \ac{DO} conditional given constant tidal height ($DO_{mtd}$, \cref{fig:do_dtd}).  Specific characteristics of the \ac{DO} time series (e.g., relative magnitude of $\epsilon_{proc}$, $DO_{adv}$, etc.) influence the ability of the model to predict unique sources of variation in observed \ac{DO}.  An extreme case is presented by time series that have no influence of tidal advection (i.e., $DO_{adv} = 0$) on the observed time series.  \Cref{fig:extreme} shows the model results from a simulated time series with no tidal component and the observed \ac{DO} signal is composed only of a diel component ($DO_{bio} = 2$).  Although the `correct' results were obtained with large window widths, the model predicts increasingly periodic contributions of the tidal ($DO_{tid}$) and mean tidal component ($DO_{mtd}$) with decreasing window widths.  These results can be explained by phase synchrony or asynchrony between the simulated tidal component and observed \ac{DO} as the model fits the time series to a non-existent tidal advection component.  As window width decreases, higher importance is given to observations within the window such that periods with high synchrony or asynchrony between observed \ac{DO} and the tidal component will have a larger influence on the model parameters.  Although application of the model to such a time series would be impractical, the results illustrate the importance of preliminary evaluations of the data and considerations for window widths.  Time series with no obvious effect of tidal advection should use relatively large window widths because results from small window widths may be impractical.

@@

The \ac{NERRS} is a federally-funded network of 28 protected estuaries established for long-term research, water-quality monitoring, education, and coastal stewardship \citep{Wenner04}.  Continuous water quality data have been collected at \ac{NERRS} sites since 1994 through the \ac{SWMP}.  In addition to providing a basis for trend evaluation, data from \ac{SWMP} provides an unprecedented opportunity to evaluate variation in water quality parameters attributed to both biological and physical processes.  Continuous \ac{SWMP} data describe \ac{DO} variation at sites with different characteristics, including wide ranges in tidal regime \citep{Sanger02} and rates of ecosystem production \citep{Caffrey03}.

Water quality from the \ac{SWMP} database \citepalias{CDMO14} were used to validate the weighted regression model beyond simulated time series.  Continous monitoring data from January 1\textsuperscript{st} to December 31\textsuperscript{st} 2012 were obtained for four stations representing a range of geographic locations (\cref{fig:case_map,tab:case_att}).  Each station was chosen specifically using measured correlations between \ac{DO} and tidal changes, suggesting strong influences of physical processes as potential confounding factors in biological \ac{DO} signals.  The four sites included Joe Leary estuary station at Padilla Bay (Washington, 48.52$^{\circ}$N 122.48$^{\circ}$W), Middle Blackwater River station at Rookery Bay (Florida, 25.93$^{\circ}$N 81.60$^{\circ}$W), Dean Creek station at Sapelo Island (Georgia, 31.39$^{\circ}$N 81.28$^{\circ}$W), and Boca Rio station at Tijuana River (California, 32.56$^{\circ}$N 117.13$^{\circ}$W).

@@

The weighted regression method was applied to the time series for each station to obtain a detided \ac{DO} estimate for estimating metabolism.  A window width of twenty days was used as a median value between the extremes evaluated in the simulations. Unlike the simulated data, the true biological \ac{DO} signal was unknown for the case studies.  Accordingly, results were evaluated based on differences between the observed and detided \ac{DO} time series, as well as correlations of \ac{DO} and metabolism estimates with tidal height before and after application of the model.  Astronomical tidal heights were predicted for each site using sonde depth data and harmonic regressions (\texttt{oce} package in R, \citealt{Foreman89}, \citetalias{RDCT14}).  We hypothesized that metabolism esimates using the detided signal would contain less `anomalous' values than those from the observed \ac{DO} time series, where `anomalous' was defined as negative production estimates during the day and positive respiration estimates during the night.  Although anomalies could be caused by processes other than tidal advection, e.g., abiotic dark oxygen production \citep{Pamatmat97}, we assume that physical processes are the dominant sources of these values.

Ecosystem metabolism was estimated using the open-water technique \citep{Odum56} as described in \citet{Caffrey13}.  The method is used to infer net ecosystem metabolism from \ac{DO} time series using the mass balance equation:
\begin{equation}
\frac{\delta DO}{\delta t} = P - R + D
\end{equation}
where the change in \ac{DO} concentration ($\delta DO$, mmol O$_2$ m$^{-3}$) over time ($\delta t$, hours) is equal to photosynthetic rate ($P$, mmol O$_2$ m$^{-3}$ hr$^{-1}$) minus respiration rate ($R$, mmol O$_2$ m$^{-3}$ hr$^{-1}$) corrected for the rate of air-sea gas exchange at the interface ($D$, mmol O$_2$ m$^{-3}$ hr$^{-1}$) \citep{Caffrey13}. $D$ is estimated as the difference between the \ac{DO} saturation concentration and observed \ac{DO}, multiplied by a volumetric reaeration coefficient, $k_a$ \citep{Thebault08}.  The diffusion-corrected \ac{DO} flux estimates were averaged during day and night for each 24 hour period in the time series, where flux is an hourly rate of \ac{DO} change as the difference between observations at time $t$ and $t+1$.  Areal respiration rates were assumed constant during the night and substracted from daily gross production estimates to yield net ecosystem metabolism (\cref{tab:case_att}).

The effects of detiding the \ac{DO} time series for each case study varied for \ac{DO} observations and metabolism estimates (\cref{tab:cor_res}).  Correlations of observed \ac{DO} time series with predicted tidal height were highly significant, with all sites indicating positive relationships except Padilla Bay where tidal increases were associated with declines in \ac{DO}  concentration.  Metabolism estimates using observed \ac{DO} were compared with daily tidal ranges at each site since the estimates represent daily integrated values.  Metabolism estimates from observed \ac{DO} were only correlated for gross production at Elkhorn Slouth and net ecosystem metabolism for Rookery Bay (\cref{tab:cor_res}).  The detided \ac{DO} time series had no significant correlations with tidal height change, whereas trends were unclear when evaluating correlations of daily tidal range with metabolism estimates based on the detided \ac{DO} time series.  For example, correlations with daily tidal range for metabolism estimated with detided time series were either unchanged (e.g., net metabolism for Rookery Bay), reduced (e.g., production for Padilla Bay), or increased (e.g., respiration for Rookery Bay) from those using observed \ac{DO}.  Instantaneous \ac{DO} flux estimates (corrected for air-sea gas exchange) form the basis of metabolism estimates and were also evaluated for correlations to tidal changes before and after detiding the observed \ac{DO} signals.  Similar to \ac{DO}, correlations of \ac{DO} flux were substantially reduced although relationships were still significant (\cref{tab:cor_res}).

The effect of detiding \ac{DO} time series varied for each case study (\cref{tab:case_res}).  The percent of daily integrated metabolism estimates that were anomalous (negative production, positive respiration) before detiding were higher for Elkhorn Slouth and Padilla Bay as compared ot Rookery Bay and Sapelo Island.  The percent of anomalous metabolism values after detiding changed for each site, with increases for Elkhorn Slough and Padilla  Bay, and decreases for Rookery Bay and Sapelo Island.  Increases were largest for Elkhorn slough such that percent anomalous estimates for productiona and respiration increased by approximately 50\%.  Reductions were largest for Sapelo Island such that the percent anomalous values for production were decreased by approximately 50\%.  Metabolism estimates using detided \ac{DO} also had decreased mean production and respiration (i.e., increasing trend towards more balanced metabolism) for Elkhorn Slough  and Padilla Bay, whereas mean production and respiration estimates were generally unchanged for Rookery Bay and Sapelo Island.  All case studies had less heterotrophic net metabolism estimates (i.e., less negative) after detiding, in addition to decreases in the standard error for all metabolism estimates.

An example from Sapelo Island further highlights the effects of weighted regression on metabolism estimates (\cref{fig:case_ex1}).  Tidal predictions for Sapelo Island indicated that the site is strongly semidiurnal with approximately two tidal peaks per 24 hour period.  Tidal ffects on the observed \ac{DO} time series were apparent such that semidiurnal variation is closely correlated to tidal height variation.  Weighted regression was successful in removing the variation in the observed \ac{DO} time series from changes in tidal height.  The detided \ac{DO} time series exhibited more consistent diel variation with photoperiod (i.e., one peak per 24 hours) as compared to the observed time series (i.e., two peaks per 24 hours).  However, metabolism before and after detiding for the period of observation was similar, suggesting that detiding the \ac{DO} signal does not have a large effect on mean daily integerated metabolism estimates.

>>>>

Multiple time series were simulated by varying the conditions in \cref{do_obs,do_bio,do_unc,do_obs_all,do_sin,do_unc_n,deltdo,deltx,do_advp,do_adv}. Specifically, the simulated data varied in the relative amount of noise in the measurement, relative amplitude of the diel \ac{DO} component, degree of association of the tide with the \ac{DO} signal, and tidal type as diurnal, semidiurnal, and mixed semidiurnal.  Three levels were evaluated for each variable: relative noise as 0, 1, and 2 standard deviations for both process and observation uncertainty, amplitude of diel biological \ac{DO} as 0, 1, and 2 mg L$^{-1}$, and \ac{DO} change from tidal advection as 0, 1, and 2 mg L$^{-1}$.  Three tidal categories were created from \cref{do_sin} using a period of 24.82 hours (principal lunar) for diurnal, 12.42 hours (principal lunar semidiurnal) for semidiurnal, and adding both diurnal and semidiurnal series for mixed semidiurnal. A total of 243 time series were created based on 81 unique combinations of parameters for each tidal category (\cref{fig:sim_ex}).  Three window widths for each variable used to estimate the total weight vector were evaluated (as half window widths): decimal time as 1, 3, and 8 days, time of day as 6, 12, and 24 hours, and tidal height as 0.25, 0.5, and 1 as a proportion of the observed value at the center of the window.  In total, 27 window width combinations were evaluated for each of 243 simulated time series, producing results for 6561 weighted regressions.

The detided or normalized values for each regression were compared to the simulated data to evaluate the ability of weighted regression to reproduce the biological \ac{DO} signals. Results were summarized using Pearson correlation coefficients and the \ac{RMSE} between the predicted and observed \ac{DO} values and the detided and biological \ac{DO} values.  Overall, the weighted regressions provided accurate results for the detided time series compared to the `true' biological time series regardless of the simulation parameters (\cref{tab:dtd_perf1}) or window widths (\cref{tab:dtd_perf2}).  Mean correlation for all time series and window widths between the detided and biological values was \Sexpr{form_fun(perf_means['cor'])}, with values ranging from \Sexpr{form_fun(perf_rngs[1,'cor'])} to \Sexpr{form_fun(perf_rngs[2, 'cor'])}.  Mean error was \Sexpr{form_fun(perf_means['err'])}, with values ranging from \Sexpr{form_fun(perf_rngs[1,'err'])} to \Sexpr{form_fun(perf_rngs[2, 'err'])}.  Simulations with very poor performance (e.g., negative correlations) were those that had minimum widths for day windows and maximum widths for hour windows, or were those with the \ac{DO} signal composed entirely of noise from observation uncertainty. Conversely, simulations with detided time series that were identical to the true time series (e.g., correlation of one, \ac{RMSE} of zero) were those for which there was no biological or tidal influence.  While the latter examples do not represent real-world scenarios, they were included in the simulations to provide verification that the weighted regressoin provide reasonable results in extreme scenarios.

Characteristics of \ac{DO} time series that contributed to improved model performance were increasing amplitude of the diel \ac{DO} component ($DO_{die}$) and increasing process uncertainty ($e_{pro}$), whereas increasing observation uncertainty contributed to decreasing performance (\cref{tab:dtd_perf1}).  Model performance was minimally influenced by magnitude of the tidal advection component ($DO_{adv}$), although performance decreased slightly with increasing tidal effects.  Increasing widths for day and tidal proportion windows contributed to increasing model performance, whereas the opposite was true for increasing hour windows (\cref{tab:dtd_perf2}).  Graphical summaries of model performance by simulation parameters (\cref{fig:err_surf1}) and half window widths (\cref{fig:err_surf2}) supports the general trends described by \cref{tab:dtd_perf1,tab:dtd_perf2}.  Scale differences between \cref{fig:err_surf1} and \cref{fig:err_surf2} emphasize that model performance was more affected by characteristics of the \ac{DO} time series rather than the selected window widths.  For example, the range of correlation values comparing the effects of half window widths (averaged across all simulation parameters, \cref{fig:err_surf2}) were approximately half the range of correlations for comparing the effects of simulation parameters (averaged across all half window widths, \cref{fig:err_surf1}).

@@

The \ac{NERRS} is a federally-funded network of 28 protected estuaries established for long-term research, water-quality monitoring, education, and coastal stewardship \citep{Wenner04}.  Continuous water quality data have been collected at \ac{NERRS} sites since 1994 through the \ac{SWMP}.  In addition to providing a basis for trend evaluation, data from \ac{SWMP} provides an ideal opportunity to evaluate long-term variation in water quality parameters attributed to both biological and physical processes.  Continuous \ac{SWMP} data can be used to describe \ac{DO} variation at sites with different characteristics, including variation from ranges in tidal regime \citep{Sanger02} and rates of ecosystem production \citep{Caffrey03,Caffrey04}.

Continuous \ac{DO} time series and tidal height measurements at four sites from the \ac{SWMP} database \citepalias{CDMO14} were used to further validate the weighted regression model.  Monitoring data from January 1\textsuperscript{st} to December 31\textsuperscript{st} 2012 were obtained from a range of geographic locations (\cref{fig:case_map,tab:case_att}). Astronomical tidal heights were predicted for each site using harmonic regression applied to the sonde depth data (\texttt{oce} package in R, \citealt{Foreman89}, \citetalias{RDCT14}). Although, the depth data represent tidal height variation from both astronomical (i.e., gravitational effects) and meteorological (e.g., wind, precipitation inflows) sources, we isolated the former given that daily metabolism estimates were more likely to be affected by repeated diel cycling from normal tidal changes.  Each station was also chosen based on high correlations between \ac{DO} and tidal changes.  The four sites included the Vierra Mouth station at Elkhorn Slough (California, 36.81$^{\circ}$N, 121.78$^{\circ}$W), Bayview Channel at Padilla Bay (Washington, 48.50$^{\circ}$N 122.50$^{\circ}$W), Middle Blackwater River station at Rookery Bay (Florida, 25.93$^{\circ}$N 81.60$^{\circ}$W), and Dean Creek station at Sapelo Island (Georgia, 31.39$^{\circ}$N 81.28$^{\circ}$W).  The stations are generally macrotidal semidiurnal or mixed semidiurnal and net heterotrophic on an annual basis.  Net heterotrophy (i.e., respiration exceeding production) is typical for shallow water systems at temperate latitudes \citep{Caffrey03}, although values in \cref{tab:case_att} are from observed \ac{DO} time series that are strongly influenced by tidal advection.

@@

The weighted regression method was applied to the time series for each station to obtain a detided \ac{DO} time series for estimating metabolism.  The half window widths that produced the most robust results for the simulations were used for the case studies: eight days, six hours, and a tidal proportion of one. Unlike the simulated data, the true biological \ac{DO} signal was unknown for the case studies.  Accordingly, results were evaluated  using correlations of \ac{DO} and metabolism estimates with tidal height before and after application of the model.  Results were also evaluated based on the occurrence of `anomalous' daily production or respiration estimates, where anomalous was defined as negative production during the day and positive respiration estimates during the night.  Anomalous values have been previously attributed to the effects of physical processes on \ac{DO} time series \citep{Caffrey03}.  We hypothesized that metabolism esimates using the detided signal would contain less `anomalous' values than those from the observed \ac{DO} time series.  Although anomalies could be caused by processes other than tidal advection, e.g., abiotic dark oxygen production \citep{Pamatmat97}, we assume that physical processes are the dominant sources of these values.

Ecosystem metabolism was estimated from the \ac{DO} time series using the open-water technique \citep{Odum56} as described in \citet{Caffrey13}.  The method is used to infer net ecosystem metabolism using the mass balance equation:
\begin{equation}
\frac{\delta DO}{\delta t} = P - R + D
\end{equation}
where the change in \ac{DO} concentration ($\delta DO$, g O$_2$ m$^{-3}$) over time ($\delta t$, hours) is equal to photosynthetic rate ($P$, g O$_2$ m$^{-3}$ hr$^{-1}$) minus respiration rate ($R$, g O$_2$ m$^{-3}$ hr$^{-1}$), corrected for air-sea gas exchange at the interface ($D$, g O$_2$ m$^{-3}$ hr$^{-1}$) \citep{Caffrey13}. $D$ is estimated as the difference between the \ac{DO} saturation concentration and observed \ac{DO}, multiplied by a volumetric reaeration coefficient, $k_a$ \citep{Thebault08}.  The diffusion-corrected \ac{DO} flux estimates were averaged during day and night for each 24 hour period in the time series, where flux is an hourly rate of \ac{DO} change as the difference between observations at time $t$ and $t+1$.  Areal respiration rates were assumed constant during the night and substracted from daily gross production estimates to yield net ecosystem metabolism (\cref{tab:case_att}).

Detiding had significant effects on the correlations between tidal height changes, \ac{DO} time series, and metabolism estimates (\cref{tab:cor_res}).  Correlations of observed \ac{DO} time series with predicted tidal height were highly significant, with all sites indicating positive relationships, except Padilla Bay where tidal increases were associated with declines in \ac{DO}  concentration.  This suggests that seaward water masses were less anoxic than landward masses, with the opposite being true for Padilla Bay. The detided \ac{DO} time series had greatly reduced correlations with tidal height, although relationships were still significant after detiding likely because of the large sample size for each site (n $\approx$ 17500). Metabolism estimates before and after detiding were compared to the mean rate of tidal height change (i.e., first derivative of the predicted tidal height) for each day during separate solar periods.  Production rates were compared to mean rates of tidal height change during the day, respiration rates were compared to mean rates of change during the night, and net metabolism rates were compared to mean rates of change for the total 24 hour period each day.  Comparison of metabolic rates to tidal changes before and after detiding produced inconsistent results (\cref{tab:cor_res}).  Correlations for Elkhorn Slough and Sapelo Island showed consistent reductions in all three metabolims estimates after detiding.  Correlations for Padilla Bay and Rookery Bay were of opposite sign and greater magnitude after detiding for production and respiration, although net metabolism estimates had reduced correlations.

The percent of daily integrated metabolism estimates that were anomalous (negative production, positive respiration) were significantly reduced for all sites after detiding (\cref{tab:case_res}).  Before detiding, anomalous values ranged from \Sexpr{form_fun(min(anom_obs[anom_obs$Metab == 'Pg', 'Value']))} (Rookery Bay) to \Sexpr{form_fun(max(anom_obs[anom_obs$Metab == 'Pg', 'Value']))} (Padilla Bay) for production and \Sexpr{form_fun(min(anom_obs[anom_obs$Metab == 'Rt', 'Value']))} (Rookery Bay) to \Sexpr{form_fun(max(anom_obs[anom_obs$Metab == 'Rt', 'Value']))} (Elkhorn Slough) for respiration as proportions of the mean annual values. Anomalous values were reduced to zero for Rookery Bay and Sapelo Island, close to zero for Padilla Bay (\Sexpr{form_fun(anom_dtd[anom_dtd$Metab == 'Pg' & anom_dtd$.id == 'Padilla Bay', 'Value'])} for production, \Sexpr{form_fun(anom_dtd[anom_dtd$Metab == 'Rt' & anom_dtd$.id == 'Padilla Bay', 'Value'])} for respiration), and reduced to approximately half the original proportions for Elkhorn Slough (\Sexpr{form_fun(anom_dtd[anom_dtd$Metab == 'Pg' & anom_dtd$.id == 'Elkhorn Slough', 'Value'])} for production, \Sexpr{form_fun(anom_dtd[anom_dtd$Metab == 'Rt' & anom_dtd$.id == 'Elkhorn Slough', 'Value'])} for respiration).  Metabolism estimates using detided \ac{DO} time series also had decreased mean production and respiration for Elkhorn Slough, increased mean production and respiration for Padilla Bay, and generally unchanged mean production and respiration for Rookery Bay and Sapelo Island (\cref{tab:case_res}).  Mean net ecosystem metabolism was unchanged for all sites.  Decreases in the standard erorr for all metabolism estimates (production, respiration, and net) were observed for all case studies after detiding.

An example from Rookery Bay further highlights the effects of weighted regression on metabolism estimates (\cref{fig:case_ex1}).  Tidal predictions for Rookery Island indicated that the site is macrotidal with mixed semidiurnal frequencies.  The effects of tidal height changes on the observed \ac{DO} time series are apparent such that semidiurnal variation is closely correlated to  the tide.  Weighted regression was successful in removing the variation in the observed \ac{DO} time series from changes in tidal height.  The detided \ac{DO} time series exhibited more consistent diel variation with photoperiod (i.e., one peak per 24 hours) as compared to the observed time series (i.e., two peaks per 24 hours).  Metabolism estimates were also largely influenced by tidal variation such that periodicity associated with spring/neap tides was apparent, with anomalous esimates for both production and respiration observed.  Metabolism estimates from the detided \ac{DO} time series did not exhibit any apparent variation from the tide.  similar trends in metabolism estimates were observed for all sites across the period of observation ().

### 2014-08-15 dep_dat.Rnw
almost ready to start discussion

@

>>>>

######
# anomalous data, from 'case_res' code chunk in 'swmp_noise.Rnw'
load('met_comp.RData')

anom_obs <- met_comp[met_comp$Metric %in% 'Anom' & met_comp$Input %in% 'Observed', ]
anom_obs$.id <- gsub('_wtreg_[0-9]*.RData$', '', anom_obs$.id)
anom_obs$.id <- factor(anom_obs$.id, levels = c('ELKVM', 'PDBBY', 'RKBMB', 'SAPDC'), labels = c('Elkhorn Slough', 'Padilla Bay', 'Rookery Bay', 'Sapelo Island'))
anom_dtd <- met_comp[met_comp$Metric %in% 'Anom' & met_comp$Input %in% 'Detided', ]
anom_dtd$.id <- gsub('_wtreg_[0-9]*.RData$', '', anom_dtd$.id)
anom_dtd$.id <- factor(anom_dtd$.id, labels = c('Elkhorn Slough', 'Padilla Bay', 'Rookery Bay', 'Sapelo Island'))
@

### 2014-08-15 discussion.Rnw
finished tables/figs for simulation assessment after detiding worked
almost ready to start discussion

air-sea gas exchange, how might this influence results
\end{comment}

>>>>

air-sea gas exchange, how might this influence results

% comment about performance of simulations, window width is a more critical issue
An examination of unique scenarios can provide insight into factors that affect the ability of weighted regression to resolve variation in observed \ac{DO} from tidal effects ($DO_{tid}$, \cref{fig:do_dtd}) and the mean response of \ac{DO} given constant tidal height ($DO_{nrm}$, \cref{fig:do_dtd}).  These unique sources of variation are represented by $DO_{bio}$ and $DO_{adv}$ as additive components of $DO_{obs}$ (\cref{do_obs}).  Although comparison of \cref{fig:err_surf1} and \cref{fig:err_surf2}) suggested that characteristics of the \ac{DO} time series had a larger influence on  model performance than window widths, examples where model fit was either very good or very poor were time series with extreme characteristics.  For example, poor performance was observed when the observation uncertainty ($\epsilon_{obs}$) was high and both process uncertainty ($\epsilon_{pro}) and tidal advection ($DO_{adv}$) were low.  These examples represent time series with excessive random variation, no auto-correlation, and no tidal influence.  Poor model performance is expected because the weighted regression attempts to isolate variation from a non-existent tidal signal in a very noisy \ac{DO} time series.  From a practical perspective, the weighted regression should not be applied to such a time series. Most of the simulations evaluated provided a reasonable reprsentation

\end{comment}
